---
title: "Untitled"
author: "Kacper Domżał"
date: "20 03 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
library(gplots)
library(ggplot2)
library(data.table)
library(devtools)
library(Biobase)
library(limma)
library(edge)
library(genefilter)
library(qvalue)
library(tidyverse)
library(data.table)
library(corpcor)
library(irlba)
library(Rtsne)
```{r dataandmodule, echo=FALSE}
library(RColorBrewer)
library(gplots)
library(ggplot2)
library(data.table)
library(devtools)
library(Biobase)
library(limma)
library(edge)
library(genefilter)
library(qvalue)
library(tidyverse)
library(data.table)
library(corpcor)
con = url("http://bowtie-bio.sourceforge.net/recount/ExpressionSets/bottomly_eset.RData")
load(file=con)
save(bottomly.eset, file="bottomly.Rdata")
load(file="bottomly.Rdata")
edata <- as.matrix(exprs(bottomly.eset))
edata <- edata[rowMeans(edata) > 10, ]
edata <- log2(as.matrix(edata) + 1)
my_palette <- colorRampPalette(c("blue", "white", "orange"))(n = 299)
```



Homework Problem 1: Make one heatmap of the aforementioned Bottomly data with 
the following options:
a) both rows and columns are clustered, 
b) show a dendrogram only on the columns., 
and c) scale in the column direction. Send only one heatmap. 
```{r problem1}
png("Domzal_problem1.png",height=2500,width=2500)
heatmap.2(edata,
          main = "Bottomly et al. Clustered", # heat map title
          notecol="black",      # change font color of cell labels to black
          density.info="none",  # turns off density plot inside color legend
          trace="none",         # turns off trace lines inside the heat map
          margins =c(12,9),     # widens margins around plot
          col=my_palette,       # use on color palette defined earlier 
          dendrogram="column",     # only draw a row dendrogram
          scale = "column",
          Rowv = TRUE,
          Colv = "Rowv")
dev.off()
```

Homework Problem 2: As shown in the plot above, the projection on the top 2 PCs doesn’t show the 
grouping by the strains. But we have many PCs to explore. Explore different combinations of PCs in scatter plots
while coloring the data points by the genetic strains. 
Find a combination of PCs that separate the strains well. Send only one scatterplot.
```{r problem2}
png("Domzal_problem2.png",height=800,width=800)
edata = t(scale(t(edata), scale=FALSE, center=TRUE))
svd.out = svd(edata)
PC.v = data.table(svd.out$v,pData(bottomly.eset))
ggplot(PC.v) + geom_point(aes(x=V3, y=V2, col=as.factor(experiment.number)))
dev.off()
```

Homework Problem 3: Make a scatter plot of the top 2 left singular vectors.
Find a combination of PCs that separate the strains well. Send only one scatterplot.
```{r problem3}
png("Domzal_problem3.png",height=700,width=700)
PC_u = data.table(svd.out$u, pData(bottomly.eset)) 
ggplot(PC_u) + geom_point(aes(x=V1, y=V2, col=as.factor(strain))) + ggtitle("Scatter plot of the top 2 left singular vectors") 
dev.off()

```


Homework Problem 4: Make one figure that contains violin plots of the top 5 left singular vectors (loadings). 
Hint/To-do: Make sure turn the top 5 left singular vectors into a data.table (or a data.frame) and 
ggplot2 to plot them altogether.
```{r problem4}
png("Domzal_problem4.png", height=700, width=700)
violin_data = as.data.frame(PC_u[, 1:5])  
violin_data_stacked = stack(violin_data)  
colnames(violin_data_stacked) = c("Val", "principle_comp")  
ggplot(violin_data_stacked, aes(x = principle_comp, y = Val)) +  
  geom_violin() + 
  ggtitle("Violin plots of the top 5 left singular vectors")
dev.off()

```

Homework Problem 5: Cluster the genes (rows) using K-means clustering (function kmeans()) 
on the original data, with k=5 clusters. Then, create a 2-dimensional t-SNE projection 
(as done previously) while using the 5 clusters to color the data points corresponding to genes.
```{r problem5}
set.seed(1)
tsne_out <- Rtsne(edata,pca=TRUE,perplexity=30)
tsne_out = data.table(tsne_out$Y)
kmeans = kmeans(x = edata, centers = 5)
edata_cluster_kmeans <- as.character(kmeans$cluster)
tsne_out$kmeans <- edata_cluster_kmeans
png("Domzal_problem5.png", height=1500, width=1500)
ggplot(tsne_out) + geom_point(aes(x=V1, y=V2, col=as.factor(kmeans)))
dev.off()

```


